Synthesis of Statically Analyzable Accelerator Networks From Sequential Programs

### 2016-04-25 paper.tex
back fill dfs

\subsection{High Level Synthesis of Accelerators}
\label{pre2}
HLS for FPGAs are being actively developed in both industry~\cite{tools:vivadohls} and academia~\cite{tools:legup}.
These HLS tools attempt to capture parallelism in a function written with high level languages like C/C++. With the help of various user provided guidance these tools generates hardware in the form of RTL. 
%Compute operations and data accesses are allocated and scheduled according to the %dependency constraints extracted from the input program, and the resource %constraints imposed by the target platform. 
The synthesized circuits usually follow a Finite State Machine with Datapath (FSMD) paradigm. Activation of a particular operator in the datapath is associated
with a certain clock cycle and the execution of the computation engine is orchestrated by a synthesized central controller.

@@

\hline
\multirow{2}{*}{}Floyd-& {Graph Algo. with Dynamic Programming}  & \multirow{2}{*}{$\approx$ 8 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Warshall & Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                                          
\hline                                                                                                           
\multirow{2}{*}{Knapsack}& Optimization with Dynamic Programming &\multirow{2}{*}{$\approx$ 5 MB}  \\
%\cline{2-3}                                                                                                                                                    
  &Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                             
\hline
\multirow{2}{*}{} &    & \multirow{2}{*}{$\approx$ 3 MB}  \\
%\cline{2-3}                                                                                                                                                    
   &         &   \\
%\cline{2-3}                                                                        
\hline            
\multirow{2}{*}{}SpMV& Matrix in Compressed Row Storage & \multirow{2}{*}{ $\approx$ 16 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Multiply &Indirect Memory Addressing      &  \\
%\cline{2-3}

@@

\begin{figure*}[htp]
\begin{center}
\includegraphics[width=0.95\linewidth]{fig/perfNew.pdf}
\caption{Performance of Conventional Accelerators and Accelerator Networks
\label{fig:acpraw}}
\end{center}
\vspace{-2.9em}
\end{figure*}

@@

With our methodology, the accelerator networks generated 
are rather competitive against the hard processor, even
without a reconfigurable cache. When the networks are directly connected to the memory through the ACP, the average performance is 2.3x that of the baseline---an 8.4x gain over the conventional accelerators. 
With the addition of caches,  
the average runtime of the accelerator networks was reduced by 18.7\%,  while that of the conventional accelerators was cut by 45.4\%. The gap between their performance is thereby reduced from 8.4x to 5.6x. This difference in improvement is due to conventional accelerators' sensitivity to the latency of data accesses, which is also manifested by its performance degradation of 40\% when the uncached HP port is used instead of ACP.

Overall, there is a significant performance advantage in the accelerator networks as our algorithm effectively addresses the weakness of conventional accelerators. If we compare the best results achieved with the
accelerator networks to that of the conventional accelerators, we see improvement
of 3.3 to 9.1 times, with an average of 5.6. 
%This performance advantage translates to energy savings as well, as illustrated in the next section.

%\subsection{Area and Energy Comparison}
%\vspace{-1.0em}
\subsection{Area comparison}
To quantify the impact of our proposed methodology on area, 
we have compared the FPGA resouce usage of conventional accelerators and
the accelerator networks. Table~\ref{tab:areacom} shows the results, 
where each implementation is complemented with two different memory subsystem configurations.
\renewcommand{\tabcolsep}{0.5pt}
\begin{table}[htbp]
%\vspace{-1.2em}
\caption{Resource Usage of Accelerators.}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | c | c | c | c | c| }
  \hline            
  \multirow{2}{*}{} &  &   \multicolumn{3}{c|}{\bf ACP  } & \multicolumn{3}{c|}{\bf ACP + 64KB Cache }   \\
 \cline{3-8} 
 {\bf Benchmark}   &    & LUT& FFs& BRAM & LUT &   FFs      & BRAM    \\
  %
  %                 & LUT &FFs&  BRAM &  LUT &   FFs      & BRAM   \\
  \hline            
  \hline

@@

\end{tabular}
\label{tab:areacom}
\end{table}

>>>>

\subsection{High Level Synthesis of Accelerators}
\label{pre2}
HLS for FPGAs are being actively developed in both industry~\cite{tools:vivadohls} and academia~\cite{tools:legup}.
These HLS tools attempt to capture parallelism in a function written with high level languages like C/C++. With the help of various user provided guidance, these tools generate hardware in the form of RTL. 
%Compute operations and data accesses are allocated and scheduled according to the %dependency constraints extracted from the input program, and the resource %constraints imposed by the target platform. 
The synthesized circuits usually follow a Finite State Machine with Datapath (FSMD) paradigm. Activation of a particular operator in the datapath is associated
with a certain clock cycle and the execution of the computation engine is orchestrated by a synthesized central controller.

@@

\hline
\multirow{2}{*}{}Floyd-& {Graph Algo. with Dynamic Programming}  & \multirow{2}{*}{$\approx$ 8 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Warshall & Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                                          
\hline                                                                                                           
\multirow{2}{*}{Knapsack}& Optimization with Dynamic Programming &\multirow{2}{*}{$\approx$ 5 MB}  \\
%\cline{2-3}                                                                                                                                                    
  &Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                             
\hline
\multirow{2}{*}{}Depth-First&  Stack-based Graph Algo. & \multirow{2}{*}{$\approx$ 3 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Search & Irregular Memory Access Pattern       &   \\
\hline            
\multirow{2}{*}{}SpMV& Matrix in Compressed Row Storage & \multirow{2}{*}{ $\approx$ 16 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Multiply &Indirect Memory Addressing      &  \\
%\cline{2-3}

@@

\begin{figure*}[htp]
\begin{center}
\includegraphics[width=0.95\linewidth]{fig/perfNew.pdf}
\vspace{-1.0em}
\caption{Performance of Conventional Accelerators and Accelerator Networks
\label{fig:acpraw}}
\end{center}
\vspace{-2.0em}
\end{figure*}

@@

With our methodology, the accelerator networks generated 
are rather competitive against the hard processor, even
without a reconfigurable cache.  For Floyd-Warshall, knapsack and SpMV multiply, when the networks are directly connected to the memory through the ACP, the average performance is 2.3x that of the baseline---an 8.4x gain over the conventional accelerators. 
With the addition of caches,  
the average runtime of the accelerator networks was reduced by 18.7\%,  while that of the conventional accelerators was cut by 45.4\%. The gap between their performance is thereby reduced from 8.4x to 5.6x. This difference in improvement is due to conventional accelerators' sensitivity to the latency of data accesses, which is also manifested by its performance degradation of 40\% when the uncached HP port is used instead of ACP.

Our partitioning algorithm also has its limitations, as demonstrated by its ineffectiveness in the depth first search benchmark.
The kernel performs very little computing but lots of memory accesses.
The use of a stack in DFS also creates a dependence cycle through the memory.
Consequently, the performance is fundamentally limited by the latency of memory access.
Thus there were only small differences between the performance of the conventional accelerator and the accelerator network, both of which achieve throughput far below that of the baseline.

Overall, for kernels suitable for FPGA acceleration, there is a significant performance advantage in the accelerator networks as our algorithm effectively addresses the weakness of conventional accelerators. If we compare the best results achieved with the
accelerator networks to that of the conventional accelerators, we see improvement
of 3.3 to 9.1 times, with an average of 5.6. 
%This performance advantage translates to energy savings as well, as illustrated in the next section.

%\subsection{Area and Energy Comparison}
%\vspace{-1.0em}
\subsection{Area comparison}
To quantify the impact of our proposed methodology on area, 
we have compared the FPGA resouce usage of conventional accelerators and
the accelerator networks. Table~\ref{tab:areacom} shows the results, 
where each implementation is complemented with two different memory subsystem configurations.
\renewcommand{\tabcolsep}{0.5pt}
\begin{table}[htbp]
\vspace{-1.5em}
\caption{Resource Usage of Accelerators.}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | c | c | c | c | c| }
  \hline            
  \multirow{2}{*}{} &  &   \multicolumn{3}{c|}{\bf ACP  } & \multicolumn{3}{c|}{\bf ACP + 64KB Cache }   \\
 \cline{3-8} 
 {\bf Benchmark}   &    & LUT& FFs& BRAM & LUT &   FFs      & BRAM    \\
  %
  %                 & LUT &FFs&  BRAM &  LUT &   FFs      & BRAM   \\
  \hline            
  \hline

@@

\end{tabular}
\vspace{-1.5em}
\label{tab:areacom}
\end{table}
